\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{cleveref} % this must be the last package to be loaded

\newcommand{\emailaddr}[1]{\href{mailto:#1}{\texttt{#1}}}

\title{\LARGE
    Hermes
}
\subtitle{Final Report for the Distributed Systems Course 2025/2026}

% Consider watching:
% https://www.youtube.com/watch?v=ihxSUsJB_14
% https://www.youtube.com/watch?v=XTFWaV55uDo

\author{
    Martin Tomassi \\ \emailaddr{martin.tomassi@studio.unibo.it}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
	Hermes is a distributed real-time messaging system designed as a feasible prototype for a university project in distributed systems. The goal of the project is not to replicate a full-scale commercial messaging platform, but to demonstrate the core principles of distributed communication through a compact yet meaningful application. The system supports authenticated one-to-one text messaging, allowing users to register, log in, and exchange messages through a desktop client. 
	
	The architecture is based on a client-server model with a distributed backend. Users authenticate through HTTP/REST, receive a JWT token, and then use a persistent WebSocket connection for real-time message exchange. Messages are handled by an event-driven backend built around Kafka, while Redis is used to support presence and routing information and PostgreSQL stores user credentials. On the client side, SQLite is used to persist local message history and synchronization metadata, enabling offline access and later reconciliation after disconnections.
	
	The project focuses on key distributed-systems concepts such as scalability, replication, fault tolerance, recovery, and eventual consistency. In particular, the system prioritizes availability over strict consistency and uses logical ordering mechanisms to preserve a meaningful message order without requiring a globally synchronized state. Hermes therefore provides a concrete example of how distributed components can cooperate to deliver a responsive, resilient, and extensible messaging service within a realistic development scope.
\end{abstract}

\section{Concept}\label{concept}

\begin{itemize}
	\item \textbf{Type of product:} The project is a real-time messaging service composed of a distributed backend and a desktop client application with a graphical user interface. The backend exposes web services and supports message exchange through a client-server architecture.
	
	\item \textbf{Use case description:} The software enables one-to-one chat between users:
	\begin{itemize}
		\item \emph{What:} Users can register and log in, then exchange instant text messages. Each message sent by a user is delivered to the intended recipient if they are online, or synchronized later if they are temporarily offline. The system does not maintain a permanent server-side conversation history; instead, message history is stored locally on the client, which also allows users to access previously received messages and compose messages while disconnected from the network.
		\item \emph{Where:} Users interact with the system from any location with an Internet connection, typically using personal devices such as laptops or desktop computers.
		\item \emph{When and how frequently:} Interaction is on-demand and frequent: users may send and receive messages continuously during a chat session. The platform supports both sporadic use and sustained conversation flow.
		\item \emph{How and which devices:} Users interact with the system through a desktop client with a graphical user interface. The client first connects to the backend through HTTP/REST to perform authentication and initial synchronization, obtaining an access token and retrieving messages received since the last access. Once this phase is completed, the client opens a persistent WebSocket connection for real-time updates.
		\item \emph{Data storage:} The system stores user-related data and message-related data. User accounts are stored in a relational database (PostgreSQL), while the messaging backbone relies on Kafka to support the exchange of messages within the distributed backend. On the client side, messages and synchronization data are stored locally in SQLite.
		\item \emph{Roles:} The main role is the end user, i.e., a chat participant. A system administrator role may exist at infrastructure level, but it is not central to the scope of this project. Standard user actions are symmetric between users.
	\end{itemize}
	
	\item \textbf{Why distribution is needed:} The system adopts a distributed architecture to satisfy the main requirements of a modern real-time messaging platform. In particular, distribution allows the system to scale more easily by adding service instances and expanding the supporting infrastructure when needed. It also improves fault tolerance and availability, since the failure of a single component does not necessarily compromise the whole service. Replication of internal components helps avoid single points of failure and supports durability of important state, while the event-driven nature of the system makes message handling more asynchronous and loosely coupled. Distribution is also useful for recovery, because it allows the system to resume operation after transient failures and to reconcile state once connectivity is restored. More generally, this architectural choice improves modularity and separation of concerns, since different parts of the system can be responsible for distinct tasks such as authentication, message propagation, presence management, and persistence. Following the CAP theorem, the system gives priority to \emph{Availability} rather than strict \emph{Consistency}. For this reason, it adopts an \emph{eventual consistency} approach: messages are accepted and propagated as soon as possible, and a consistent state is recovered later through synchronization mechanisms. To support this behavior, the system can use Lamport clocks to preserve a meaningful logical ordering of events without requiring a fully synchronized global state. This choice is appropriate for a chat application, where responsiveness is more important than enforcing strong consistency at every instant.
\end{itemize}

\section{Requirements Elicitation and Analysis}\label{requirements}

\subsection{Functional Requirements}

\begin{itemize}
	\item \textbf{FR1 -- User registration and authentication:} The system shall allow a user to create an account and authenticate securely before accessing the messaging service.
	
	\textbf{Acceptance criterion:} A user can register with valid credentials, log in successfully, and access the application only after authentication.
	
	\item \textbf{FR2 -- One-to-one messaging:} The system shall allow authenticated users to exchange private text messages with other registered users.
	
	\textbf{Acceptance criterion:} A user can send a text message to another registered user, and the recipient can read it if they are connected or later after synchronization.
	
	\item \textbf{FR3 -- Real-time message delivery:} The system shall deliver messages in real time when both users are online.
	
	\textbf{Acceptance criterion:} When two users are connected, a message sent by one user becomes visible to the recipient without requiring manual refresh.
	
	\item \textbf{FR4 -- Offline message synchronization:} The system shall allow users to recover messages that were received while they were temporarily offline.
	
	\textbf{Acceptance criterion:} After a user reconnects and logs in again, previously missed messages are retrieved and displayed in the client.
	
	\item \textbf{FR5 -- Local conversation storage:} The system shall keep a local history of messages on the client side so that users can access past conversations even without an active network connection.
	
	\textbf{Acceptance criterion:} A user can open the client and view previously stored messages from the local database.
	
	\item \textbf{FR6 -- Message ordering support:} The system shall preserve a meaningful order of messages within a conversation, even when delivery is asynchronous.
	
	\textbf{Acceptance criterion:} Messages are displayed in an order that is consistent with the logical progression of the conversation.
\end{itemize}

\subsection{Non-Functional Requirements}

\begin{itemize}
	\item \textbf{NFR1 -- Availability:} The system shall remain usable even when some internal components are temporarily unavailable.
	
	\textbf{Acceptance criterion:} A temporary failure of one internal component does not completely stop the service, and the system resumes normal operation after recovery.
	
	\item \textbf{NFR2 -- Eventual consistency:} The system shall ensure that the state observed by users converges over time after temporary disconnections or delays.
	
	\textbf{Acceptance criterion:} After synchronization, users eventually observe a consistent view of exchanged messages.
	
	\item \textbf{NFR3 -- Responsiveness:} The system shall support low-latency interaction suitable for a real-time chat application.
	
	\textbf{Acceptance criterion:} Sending and receiving a message should feel immediate from the user's perspective during normal operation.
	
	\item \textbf{NFR4 -- Reliability and recovery:} The system shall recover gracefully from transient failures and support state reconciliation after disconnections.
	
	\textbf{Acceptance criterion:} After a disconnection, the system is able to restore missing messages and reestablish a consistent conversation state.
	
	\item \textbf{NFR5 -- Scalability:} The system shall support growth in the number of users, messages, and concurrent connections without requiring a redesign of the overall architecture.
	
	\textbf{Acceptance criterion:} The system can be extended by adding more service instances or infrastructure resources while preserving the same logical behavior.
	
	\item \textbf{NFR6 -- Durability:} The system shall ensure that messages are not lost due to transient failures.
	
	\textbf{Acceptance criterion:} Messages sent by a user are eventually delivered or recoverable after temporary failures.
	
	\item \textbf{NFR7 -- Maintainability:} The system shall be structured in a modular way so that different responsibilities can evolve independently.
	
	\textbf{Acceptance criterion:} Changes to one subsystem, such as authentication or message delivery, can be introduced with limited impact on the others.
\end{itemize}

\subsection{Implementation Requirements}

\begin{itemize}
	\item \textbf{IR1 -- Limited development scope:} The system shall be designed and implemented by a single developer within a limited time budget (approximately 90 hours).
	
	\textbf{Acceptance criterion:} The final system includes a working prototype that covers the core functionalities within the given time constraints.
	
	\item \textbf{IR2 -- Local deployment environment:} The system shall be deployable and testable on a single machine using a local environment.
	
	\textbf{Acceptance criterion:} The system can be executed and tested locally without requiring external infrastructure.
	
	\item \textbf{IR3 -- Use of container orchestration:} The system shall be deployable in a containerized environment using Kubernetes for educational purposes.
	
	\textbf{Acceptance criterion:} The system can be deployed and run using a local Kubernetes setup (e.g., Minikube).
	
	\item \textbf{IR4 -- Technology familiarity constraint:} The implementation should rely on technologies that are compatible with the developer’s existing knowledge and learning goals.
	
	\textbf{Acceptance criterion:} The chosen technologies allow the system to be implemented within the project constraints and support the intended learning objectives.
\end{itemize}

\subsection{Glossary}

\begin{itemize}
	\item \textbf{JWT (JSON Web Token):} A token used to represent the authenticated identity of a user and to authorize requests to the system.
	\item \textbf{REST (Representational State Transfer):} An architectural style for designing networked applications, typically based on HTTP, used here for client-server communication during authentication and synchronization.
	\item \textbf{WebSocket:} A communication protocol that enables persistent bidirectional interaction between the client and the server for real-time message exchange.
	\item \textbf{Kafka:} A distributed event streaming platform used as the backbone of the system to propagate messages between components.
	\item \textbf{Event-driven architecture:} A design paradigm in which system components communicate through the production and consumption of events, enabling loose coupling and asynchronous interaction.
	\item \textbf{Lamport clock:} A logical clock mechanism used to maintain a consistent ordering of events in a distributed system without relying on synchronized physical time.
	\item \textbf{Eventual consistency:} A consistency model in which the system guarantees that all participants will converge to the same state over time, even if temporary inconsistencies occur.
	\item \textbf{Presence information:} Information that indicates whether a user is currently online and reachable, used to support message routing and delivery decisions.
	\item \textbf{Reconciliation:} The process of restoring a consistent system state after temporary inconsistencies or disconnections.
\end{itemize}

\subsection{Relevant Distributed System Features}
\label{ds-features}

\begin{itemize}
	\item \textbf{Transparency} is considered an important but not absolute requirement. From the user perspective, the system should hide most of the underlying distributed complexity. In particular, users should not be aware of how messages are propagated, where components are deployed, or how many internal services are involved. The interaction should resemble that of a standard messaging application, where sending and receiving messages appears simple and immediate.
	
	However, transparency is only partially required. Some aspects of distribution are intentionally visible at a conceptual level due to the educational nature of the project. For example, temporary delays, message reordering, or synchronization after disconnection may be observable by the user, reflecting the underlying distributed behavior of the system. This is consistent with the choice of prioritizing availability over strict consistency.
	
	Location transparency is not fully enforced in the current prototype, since all client requests are routed to a single back-end entry point without location-aware policies. Replication and failure transparency are also only partial: the system is designed to tolerate failures and continue operating, but transient inconsistencies or delays may still be visible during recovery.
	
	From the developer perspective, distribution is explicit rather than hidden, since distributed components, asynchronous communication, and event-based interaction are central to the learning goals of the project.
	
	\item \textbf{Fault tolerance and dependability -- availability, reliability, integrity, maintainability, safety:} The system places a strong emphasis on availability. If a component fails, the service should continue operating in a degraded but still usable state whenever possible. Temporary unavailability may delay message delivery or synchronization, but it should not prevent users from using the application entirely.
	
	Data loss is unacceptable in normal operation. Messages sent by users should eventually be delivered or made available after recovery, even in the presence of transient failures. Short-lived inconsistencies may occur because of the asynchronous nature of the system, but data integrity should be preserved in the long term.
	
	Recovery from faults should be automatic and reasonably fast, without requiring user intervention. After a failure or disconnection, the system should restore connectivity, retrieve missing messages, and re-establish a consistent state through synchronization and reconciliation.
	
	Reliability is supported by asynchronous communication and persistent state management. Maintainability is also important: the system is structured in a modular way so that individual components can be updated, replaced, or debugged independently. Safety, understood as protection against harmful behavior in a safety-critical sense, is not a primary concern for this application domain.
	
	\item \textbf{Scalability:} The system is expected to handle increasing numbers of users, messages, and concurrent connections. Even though the prototype is developed within a limited scope, it is designed with the assumption that the workload may grow over time. In particular, the system should support multiple active users exchanging messages simultaneously without requiring a redesign of the overall architecture.
	
	Scalability is therefore an important requirement, as real-time messaging systems are inherently subject to variable and potentially high loads. The system should be able to grow both in terms of the number of users and the volume of exchanged messages by extending resources and services incrementally.
	
	However, scalability is considered at a conceptual level in this project. The goal is not to achieve large-scale production performance, but to demonstrate how a distributed architecture can support growth through modularity and extensibility.
	
	\item \textbf{Security and trust:} The system processes and stores sensitive data, in particular user credentials and private messages exchanged between users. For this reason, security is an important requirement. While the prototype does not aim to provide production-grade security guarantees, it must ensure a reasonable level of protection against unauthorized access and misuse of data.
	
	The system is based on a single main user role: authenticated users who can access the messaging functionality. No complex role hierarchy or fine-grained authorization model is required within the scope of this project. However, only authenticated users should be able to interact with the system and access their own communication data.
	
	Authentication is therefore a core requirement, and the system must verify the identity of users before granting access to its services. Basic authorization is also required to ensure that users can only send and receive messages that belong to them. Trust in the system is established through controlled access and by preventing unauthorized interactions between users.
	
	More advanced security mechanisms, such as sophisticated attack prevention or comprehensive access control policies, are outside the scope of the project. The focus is instead on the fundamental mechanisms that enable secure communication in a distributed environment.
	
	\item \textbf{Resource sharing:} The system involves multiple users and distributed components that need to access shared resources. In particular, users share the messaging infrastructure to exchange messages, while back-end components share communication channels, user data, and transient state such as presence information. Resource sharing is therefore an inherent aspect of the system.
	
	Coordination is required to ensure correct behavior when accessing shared resources. For example, message exchange must be coordinated so that messages are correctly routed and eventually delivered, even in the presence of asynchronous communication and temporary disconnections. Similarly, shared state such as user presence must be handled in a way that avoids inconsistencies that could compromise the user experience.
	
	However, strict synchronization is not always enforced. In line with the overall design philosophy of the system, coordination mechanisms are lightweight and tolerate temporary inconsistencies. The system favors eventual consistency and asynchronous interaction over strong coordination.
	
	\item \textbf{Openness, interoperability, and heterogeneity of components:} The system combines several components that may rely on different technologies and communication paradigms. Even though the prototype is developed within a controlled environment, it reflects a heterogeneous architecture in which different subsystems cooperate to provide the overall functionality. This makes interoperability an important aspect, as components must communicate effectively despite differences in implementation.
	
	Standardization plays a relevant role in enabling interoperability. The system relies on well-established communication models and interfaces, allowing components to exchange information in a consistent and predictable way. This facilitates integration between different parts of the system and makes it easier to extend or replace individual components without affecting the entire architecture.
	
	Openness is considered at a conceptual level. The system is not explicitly designed to interoperate with external third-party services in its current form; however, it is structured in a way that could support future extensions or integration scenarios. The prototype is therefore open to evolution and adaptation, even if full interoperability with external systems is not a primary objective.
	
	\item \textbf{Evolvability and maintainability:} The system is expected to be updated and extended after its initial development, even though it is implemented as a prototype. In particular, the architecture is conceived to allow the addition of new features, improvements to existing functionalities, or replacement of individual components without requiring a complete redesign.
	
	Maintainability is therefore a relevant concern. The system is structured in a modular way, with clearly separated responsibilities, so that different parts can be understood, modified, and tested independently. This reduces the complexity of future changes and facilitates debugging and incremental development.
	
	Long-term maintenance is not a primary objective of the project, since the system is not intended for production use. However, adopting maintainable design principles is still important to ensure that the system remains understandable and adaptable over time.
	
	\item \textbf{Performance, concurrency, computation and communication efficiency, bandwidth:} The system is expected to provide a responsive user experience, as it targets real-time messaging scenarios. While there are no strict quantitative requirements on response time or throughput, message delivery should appear near-instantaneous under normal operating conditions.
	
	Concurrency is an important aspect of the system. Multiple users may interact with the platform simultaneously, sending and receiving messages in parallel. The system must therefore handle concurrent operations without significant degradation in responsiveness or correctness.
	
	Communication efficiency is relevant due to the distributed nature of the system. Messages are exchanged between multiple components, and network communication plays a central role in overall performance. Although bandwidth is not expected to be a critical limitation in the target environment, the system should avoid unnecessary data transfer and redundant communication whenever possible.
	
	Computation efficiency is considered at a conceptual level. The system should not introduce unnecessary processing overhead that could negatively affect responsiveness. However, low-level optimization is not a primary goal of the project.
	
	\item \textbf{Economy, costs:} The system is designed within a strict development budget and time constraint, as it is implemented by a single developer for a university project. Resource usage should be kept reasonable to fit within the limited scope and infrastructure available.
	
	Deployment and operational costs are expected to be low, since the prototype will primarily run on a local machine.
	
	While high efficiency is not the primary focus, economical design choices are encouraged to keep both development effort and operational cost manageable.
\end{itemize}

\section{Design}\label{design}

This chapter explains the strategies used to meet the requirements identified in the analysis. 
%
Ideally, the design should be the same, regardless of the technological choices made during the implementation phase.

\begin{quote}
	You can re-arrange the sections as you prefer, but all the sections must be present in the end
\end{quote}

\noindent
\textbf{Important:} try to motivate your design choices in relation to the requirements and features identified 
in \cref{requirements} and \cref{ds-features}.

\subsection{Architecture}\label{architecture}

\begin{itemize}
	\item Which architectural style?
	%
	\begin{itemize}
		\item why?
	\end{itemize}
\end{itemize}

\subsection{Infrastructure}\label{infrastructure}

\begin{itemize}
	\item are there \emph{infrastructural components} that need to be introduced? \emph{how many}?
	%
	\begin{itemize}
		\item e.g.~\emph{clients}, \emph{servers}, \emph{load balancers},
		\emph{caches}, \emph{databases}, \emph{message brokers},
		\emph{queues}, \emph{workers}, \emph{proxies}, \emph{firewalls},
		\emph{CDNs}, \emph{etc.}
	\end{itemize}
	
	\item how do components \emph{distribute} over the network? \emph{where}?
	%
	\begin{itemize}
		\item e.g.~do servers / brokers / databases / etc. sit on the same
		machine? on the same network? on the same datacenter? on the same
		continent?
	\end{itemize}
	
	\item how do components \emph{find} each other?
	%
	\begin{itemize}
		\item how to \emph{name} components?
		\item e.g.~DNS, \emph{service discovery}, \emph{load balancing},
		\emph{etc.}
	\end{itemize}
\end{itemize}

\begin{quote}
	Component diagrams are welcome here
\end{quote}

\subsection{Modelling}\label{modelling}

\begin{itemize}
	\item which \textbf{domain entities} are there?
	%
	\begin{itemize}
		\item e.g.~\emph{users}, \emph{products}, \emph{orders}, \emph{etc.}
	\end{itemize}
	
	\item how do \emph{domain entities} \textbf{map to} \emph{infrastructural components}?
	%
	\begin{itemize}
		\item e.g.~state of a video game on central server, while inputs/representations on clients
		\item e.g.~where to store messages in an IM app? for how long?
	\end{itemize}
	
	\item which \textbf{domain events} are there?
	%
	\begin{itemize}
		\item e.g.~\emph{user registered}, \emph{product added to cart}, \emph{order placed}, \emph{etc.}
	\end{itemize}
	
	\item which sorts of \textbf{messages} are exchanged?
	%
	\begin{itemize}
		\item e.g.~\emph{commands}, \emph{events}, \emph{queries}, \emph{etc.}
	\end{itemize}
	
	\item what information does the \textbf{state} of the system comprehend
	%
	\begin{itemize}
		\item e.g.~\emph{users' data}, \emph{products' data}, \emph{orders' data}, \emph{etc.}
	\end{itemize}
\end{itemize}

\begin{quote}
	Class diagram are welcome here
\end{quote}

\subsection{Interaction}\label{interaction}

\begin{itemize}
	\item how do components \emph{communicate}? \emph{when}? \emph{what}?
	\item \emph{which} \textbf{interaction patterns} do they enact?
\end{itemize}

\begin{quote}
	Sequence diagrams are welcome here
\end{quote}

\subsection{Behaviour}\label{behaviour}

\begin{itemize}
	\item how does \emph{each} component \textbf{behave} individually (e.g.~in \emph{response} to \emph{events} or messages)?
	\begin{itemize}
		\item some components may be \emph{stateful}, others \emph{stateless}
	\end{itemize}
	
	\item which components are in charge of updating the \textbf{state} of the system? \emph{when}? \emph{how}?
\end{itemize}

\begin{quote}
	State diagrams are welcome here
\end{quote}

\subsection{Data and Consistency Issues}\label{data-and-consistency-issues}

\begin{itemize}
	\item Is there any data that needs to be stored?
	%
	\begin{itemize}
		\item \emph{what} data? \emph{where}? \emph{why}?
	\end{itemize}
	
	\item how should \emph{persistent data} be \textbf{stored}?
	%
	\begin{itemize}
		\item e.g.~relations, documents, key-value, graph, etc.
		\item why?
	\end{itemize}
	
	\item Which components perform queries on the database?
	%
	\begin{itemize}
		\item \emph{when}? \emph{which} queries? \emph{why}?
		\item concurrent read? concurrent write? why?
	\end{itemize}
	
	\item Is there any data that needs to be shared between components?
	%
	\begin{itemize}
		\item \emph{why}? \emph{what} data?
	\end{itemize}
\end{itemize}

\subsection{Fault-Tolerance}\label{fault-tolerance}

\begin{itemize}
	\item Is there any form of data \textbf{replication} / federation / sharing?
	%
	\begin{itemize}
		\item \emph{why}? \emph{how} does it work?
	\end{itemize}
	
	\item Is there any \textbf{heart-beating}, \textbf{timeout}, \textbf{retry mechanism}?
	%
	\begin{itemize}
		\item \emph{why}? \emph{among} which components? \emph{how} does it work?
	\end{itemize}
	
	\item Is there any form of \textbf{error handling}?
	%
	\begin{itemize}
		\item \emph{what} happens when a component fails? \emph{why}? \emph{how}?
	\end{itemize}
\end{itemize}

\subsection{Availability}\label{availability}

\begin{itemize}
	\item Is there any \textbf{caching} mechanism?
	%
	\begin{itemize}
		\item \emph{where}? \emph{why}?
	\end{itemize}
	
	\item Is there any form of \textbf{load balancing}?
	%
	\begin{itemize}
		\item \emph{where}? \emph{why}?
	\end{itemize}
	
	\item In case of \textbf{network partitioning}, how does the system behave?
	%
	\begin{itemize}
		\item \emph{why}? \emph{how}?
	\end{itemize}
\end{itemize}

\subsection{Security}\label{security}

\begin{itemize}
	\item Is there any form of \textbf{authentication}?
	%
	\begin{itemize}
		\item \emph{where}? \emph{why}?
	\end{itemize}
	
	\item Is there any form of \textbf{authorization}?
	%
	\begin{itemize}
		\item which sort of \emph{access control}?
		\item which sorts of users / \emph{roles}? which \emph{access rights}?
	\end{itemize}
	
	\item Are \textbf{cryptographic schemas} being used?
	%
	\begin{itemize}
		\item e.g.~token verification,
		\item e.g.~data encryption, etc.
	\end{itemize}
\end{itemize}

\section{Implementation}\label{implementation}

Please report here all the implementation (technology-dependent) choices you made while implementing your design.

\begin{quote}
	If you run out of time for this project, you may consider leaving some aspect unimplemented, 
	and simply discuss how you would have implemented them.
	%
	In this case, better would be to discuss unimplemented features in \cref{future-works}.
\end{quote}

\begin{itemize}
	\item which \textbf{network protocols} to use?
	%
	\begin{itemize}
		\item e.g.~UDP, TCP, HTTP, WebSockets, gRPC, XMPP, AMQP, MQTT, etc.
	\end{itemize}
	
	\item how should \emph{in-transit data} be \textbf{represented}?
	%
	\begin{itemize}
		\item e.g.~JSON, XML, YAML, Protocol Buffers, etc.
	\end{itemize}
	
	\item how should \emph{databases} be \textbf{queried}?
	%
	\begin{itemize}
		\item e.g.~SQL, NoSQL, etc.
	\end{itemize}
	
	\item how should components be \emph{authenticated}?
	%
	\begin{itemize}
		\item e.g.~OAuth, JWT, etc.
	\end{itemize}
	
	\item how should components be \emph{authorized}?
	%
	\begin{itemize}
		\item e.g.~RBAC, ABAC, etc.
	\end{itemize}
\end{itemize}

\subsection{Technological details}\label{technological-details}

\begin{itemize}
	\item any particular \emph{framework} / \emph{technology} being exploited goes here
\end{itemize}

\section{Validation}\label{validation}

\subsection{Automatic Testing}\label{automatic-testing}

\begin{itemize}
	\item how were individual components \textbf{\emph{unit}-test}ed?
	
	\item how was communication, interaction, and/or integration among components tested?
	
	\item how to \textbf{\emph{end-to-end}-test} the system?
	%
	\begin{itemize}
		\item e.g.~production vs.~test environment
	\end{itemize}
	
	\item for each test specify:
	%
	\begin{itemize}
		\item rationale of individual tests
		\item how were the test automated
		\item how to run them
		\item which requirement they are testing, if any
	\end{itemize}
\end{itemize}

\begin{quote}
	recall that \emph{deployment} \textbf{automation} is commonly used to \emph{test} the system in \emph{production-like} environment
\end{quote}

\begin{quote}
	recall to test corner cases (crashes, errors, etc.)
\end{quote}

\subsection{Acceptance test}\label{acceptance-test}

\begin{itemize}
	\item did you perform any \emph{manual} testing?
	%
	\begin{itemize}
		\item what did you test?
		\item why wasn't it automatic?
	\end{itemize}
\end{itemize}

\section{Deployment}\label{deployment}

\begin{itemize}
	\item should one install your software from scratch, how to do it?
	%
	\begin{itemize}
		\item provide instructions
		\item provide expected outcomes
	\end{itemize}
	
	\item what software should be installed on the machines to run your project? which versions?
	
	\item should one set environment variables or configuration files?
	
	\item if you're using containerization (e.g.~Docker), 
	describe how you engineered your deployment scrips (e.g. \texttt{docker-compose.yml} files)
\end{itemize}

\section{User Guide}\label{user-guide}

\begin{itemize}
	\item how to use your software?
	%
	\begin{itemize}
		\item provide instructions
		\item provide expected outcomes
		\item provide screenshots if possible
	\end{itemize}
\end{itemize}

\section{Self-evaluation}\label{self-evaluation}

\begin{itemize}
	\item An individual section is required for each member of the group
	\item Each member must self-evaluate their work, listing the strengths and weaknesses of the product
	\item Each member must describe their role within the group as objectively as possible.
\end{itemize}

It should be noted that each student is only responsible for their own section.

\section{Future works}\label{future-works}

Discuss possible future works, improvements, extensions, optimizations, etc.

Also discuss here any unimplemented feature, and how you would have implemented it.

\nocite{*}
\bibliographystyle{plain}
\bibliography{references}

\end{document}